PANNEURO_CFS_cleaning_WHO_pres

### 2020-10-26 PANNEURO_analysis_WHO_pres.Rmd
Fixed NBR IDs (hopefully).

# ALSfrs
```{ALSfrs col names r}
als.col.names <- dat.raw %>%
  select(
    starts_with("alsfrs") & ends_with("numeric")
  ) %>%
  colnames()

@@

```{ALSfrs row sums r}
dat.raw <- dat.raw %>%
  mutate(als.sum.score =
  rowSums(select(.,als.col.names), na.rm = F)
   )
```

```{ALSfrs frequencies r}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_probable_case_incl_screening)) %>%
  filter(!is.na(als.sum.score)) %>%
  group_by(covid_probable_case_incl_screening) %>%
  count(covid_probable_case_incl_screening, als.sum.score) %>%
  mutate(Prop = round(n/sum(n), 2))

```
# ESS
```{ESS col names r}
ess.col.names <- dat.raw %>%
  select(
    starts_with("ess") & ends_with("numeric")
  ) %>%
  colnames()

@@

```{ESS row sums r}
dat.raw <- dat.raw %>%
  mutate(ess.col.names =
   rowSums(select(.,ess.col.names), na.rm = F)
   )
```

```{ESS frequencies r}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_probable_case_incl_screening)) %>%
  filter(!is.na(ess.col.names)) %>%
  group_by(covid_probable_case_incl_screening) %>%
  count(covid_probable_case_incl_screening, ess.col.names) %>%
  mutate(Prop = round(n/sum(n), 2))
```

@@

```{r}
cfs.col.names <- dat.raw %>%
  select(
    starts_with("cfs") & ends_with("numeric")
  ) %>%
  colnames()
cfs.col.names
```
```{r}
dat.raw <- dat.raw %>%
  mutate(cfs.sum_score =
   rowSums(select(.,cfs.col.names), na.rm = F) - 11
   )
cfs.col.names.physio <- cfs.col.names[1:7]

>>>>

# ALSfrs
```{r ALSfrs col names}
als.col.names <- dat.raw %>%
  select(
    starts_with("alsfrs") & ends_with("numeric")
  ) %>%
  colnames()

@@

```{r ALSfrs row sums}
dat.raw <- dat.raw %>%
  mutate(als.sum.score =
  rowSums(select(.,als.col.names), na.rm = F)
   )
```

```{r ALSfrs frequencies}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_probable_case_incl_screening)) %>%
  filter(!is.na(als.sum.score)) %>%
  group_by(covid_probable_case_incl_screening) %>%
  count(covid_probable_case_incl_screening, als.sum.score) %>%
  mutate(Prop = round(n/sum(n), 2))

```
# ESS
```{r ESS col names}
ess.col.names <- dat.raw %>%
  select(
    starts_with("ess") & ends_with("numeric")
  ) %>%
  colnames()

@@

```{r ESS row sums}
dat.raw <- dat.raw %>%
  mutate(ess.col.names =
   rowSums(select(.,ess.col.names), na.rm = F)
   )
```

```{r ESS frequencies}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_probable_case_incl_screening)) %>%
  filter(!is.na(ess.col.names)) %>%
  group_by(covid_probable_case_incl_screening) %>%
  count(covid_probable_case_incl_screening, ess.col.names) %>%
  mutate(Prop = round(n/sum(n), 2))
```

@@

```{r select cfs variables}
cfs.col.names <- dat.raw %>%
  select(
    starts_with("cfs") & ends_with("numeric")
  ) %>%
  colnames()
cfs.col.names
```

```{r calculate sum score}
dat.raw <- dat.raw %>%
  mutate(cfs.sum_score =
   rowSums(select(.,cfs.col.names), na.rm = F) - 11
   )
cfs.col.names.physio <- cfs.col.names[1:7]

### 2020-10-26 PANNEURO_preprocessing_CFS_WHO-pres.Rmd
Fixed NBR IDs (hopefully).

# NBR cohort
nbr.cfs.raw.id <- cfs.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  #distinct(externalDataReference, .keep_all = TRUE) %>% # DO NOT remove duplicates based on ID. We will need to divide them by startDate. 
  filter(startsWith(externalDataReference, 'SP', trim=TRUE)) %>%  #NBR cohort
  separate(externalDataReference, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>% 
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate, #Date when participant completed set B
         cfs.do_you_have_problems_with_tiredness = NA.8.do_you_have_problems_with_tiredness, 
         cfs.do_you_need_to_rest_more = NA.8.do_you_need_to_rest_more, 
         cfs.do_you_feel_sleepy_or_drowsy = NA.8.do_you_feel_sleepy_or_drowsy, 
         cfs.do_you_have_problems_starting_things = NA.8.do_you_have_problems_starting_things, 
         cfs.do_you_lack_energy = NA.8.do_you_lack_energy, 
         cfs.strength_muscles = NA.8.strength_muscles,
         cfs.do_you_feel_weak = NA.8.do_you_feel_weak,
         cfs.do_you_have_difficulty_concentrating = NA.8.do_you_have_difficulty_concentrating,
         cfs.make_slips_speaking_tongue = NA.8.make_slips_speaking_tongue,
         cfs.correct_word_find_difficult = NA.8.correct_word_find_difficult,
         cfs.how_is_your_memory = NA.8.how_is_your_memory
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

#GLAD/EDGI cohorts
glad.edgi.cfs.raw.id <- cfs.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  #distinct(externalDataReference, .keep_all = TRUE) %>% # DO NOT remove duplicates based on ID. We will need to divide them by startDate. 
  filter(startsWith(externalDataReference, 'GLAD', trim=TRUE) | startsWith(externalDataReference, 'EDGI', trim=TRUE)) %>% 
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate, #Date when participant completed set B
         cfs.do_you_have_problems_with_tiredness = NA.8.do_you_have_problems_with_tiredness, 
         cfs.do_you_need_to_rest_more = NA.8.do_you_need_to_rest_more, 
         cfs.do_you_feel_sleepy_or_drowsy = NA.8.do_you_feel_sleepy_or_drowsy, 
         cfs.do_you_have_problems_starting_things = NA.8.do_you_have_problems_starting_things, 
         cfs.do_you_lack_energy = NA.8.do_you_lack_energy, 
         cfs.strength_muscles = NA.8.strength_muscles,
         cfs.do_you_feel_weak = NA.8.do_you_feel_weak,
         cfs.do_you_have_difficulty_concentrating = NA.8.do_you_have_difficulty_concentrating,
         cfs.make_slips_speaking_tongue = NA.8.make_slips_speaking_tongue,
         cfs.correct_word_find_difficult = NA.8.correct_word_find_difficult,
         cfs.how_is_your_memory = NA.8.how_is_your_memory
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Merge datasets
cfs.raw.id <- bind_rows(
  nbr.cfs.raw.id,
  glad.edgi.cfs.raw.id
  )

@@

>>>>

# NBR cohort
nbr.cfs.raw.id <- cfs.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  #distinct(externalDataReference, .keep_all = TRUE) %>% # DO NOT remove duplicates based on ID. We will need to divide them by startDate. 
  filter(str_detect(externalDataReference, "SP")) %>%  #NBR cohort
  separate(externalDataReference, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>% 
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate, #Date when participant completed set B
         cfs.do_you_have_problems_with_tiredness = NA.8.do_you_have_problems_with_tiredness, 
         cfs.do_you_need_to_rest_more = NA.8.do_you_need_to_rest_more, 
         cfs.do_you_feel_sleepy_or_drowsy = NA.8.do_you_feel_sleepy_or_drowsy, 
         cfs.do_you_have_problems_starting_things = NA.8.do_you_have_problems_starting_things, 
         cfs.do_you_lack_energy = NA.8.do_you_lack_energy, 
         cfs.strength_muscles = NA.8.strength_muscles,
         cfs.do_you_feel_weak = NA.8.do_you_feel_weak,
         cfs.do_you_have_difficulty_concentrating = NA.8.do_you_have_difficulty_concentrating,
         cfs.make_slips_speaking_tongue = NA.8.make_slips_speaking_tongue,
         cfs.correct_word_find_difficult = NA.8.correct_word_find_difficult,
         cfs.how_is_your_memory = NA.8.how_is_your_memory
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

#GLAD/EDGI cohorts
glad.edgi.cfs.raw.id <- cfs.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  #distinct(externalDataReference, .keep_all = TRUE) %>% # DO NOT remove duplicates based on ID. We will need to divide them by startDate. 
  filter(str_detect(externalDataReference, "GLAD") | str_detect(externalDataReference, "EDGI")) %>% 
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate, #Date when participant completed set B
         cfs.do_you_have_problems_with_tiredness = NA.8.do_you_have_problems_with_tiredness, 
         cfs.do_you_need_to_rest_more = NA.8.do_you_need_to_rest_more, 
         cfs.do_you_feel_sleepy_or_drowsy = NA.8.do_you_feel_sleepy_or_drowsy, 
         cfs.do_you_have_problems_starting_things = NA.8.do_you_have_problems_starting_things, 
         cfs.do_you_lack_energy = NA.8.do_you_lack_energy, 
         cfs.strength_muscles = NA.8.strength_muscles,
         cfs.do_you_feel_weak = NA.8.do_you_feel_weak,
         cfs.do_you_have_difficulty_concentrating = NA.8.do_you_have_difficulty_concentrating,
         cfs.make_slips_speaking_tongue = NA.8.make_slips_speaking_tongue,
         cfs.correct_word_find_difficult = NA.8.correct_word_find_difficult,
         cfs.how_is_your_memory = NA.8.how_is_your_memory
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r rename NBR cohort}
nbr.cfs.raw.id <- nbr.cfs.raw.id%>%
  mutate(
    Sample = 
    recode(
      Sample, 
      "SP0068" = "NBR"
    )
  )
```

```{r merge datasets}
# Merge datasets
cfs.raw.id <- bind_rows(
  nbr.cfs.raw.id,
  glad.edgi.cfs.raw.id
  )
